---
title: "Retrieving Western Movies with TMDB API"
author: "Ahmet Yiğit Doğan"
date: "14 Dec, 2022"
output:
  pdf_document:
    toc: yes
    toc_depth: 1
    number_sections: yes
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.align="center")
```

\newpage

# Introduction

In this study, an API call will be sent to the The Movie Database (TMDB) servers to retrieve a list of Italian Western Movies. The resulting table will include th first 20 movies with the highest number of user votes. 

One can access to the API details by clicking [here](https://developers.themoviedb.org/3/getting-started). The API provides a variety of features to get movies and TV Shows related data such as cast, synopsis, user votes, and country based statistics.

The steps taken during the preparation of this report can be summarized as follows:

- Receiving an API Key by following the 4 steps addressed [here](https://developers.themoviedb.org/3/getting-started/introduction).

- Setting up an R Markdown file and preparing the environment by installing the related packages.

- Performing test calls to check whether the environment is working.

- Calling the base URL's and extending them with additional queries to obtain more specific results.

- The main problem with the code was the readability due to the length of the URL's. Thus, the URL's are decomposed into three parts (body, query strings, and api key) by using the base R function "gsub" along with a custom function called "extend_URL". This way, the code became more clean and also reusable with different API keys and queries.

```{r intro}

# Installing the required R packages
# install.packages("httr")
# install.packages("jsonlite")

# Library imports

library("httr")
library("jsonlite")

# Setting the API key

api_key <- "5169379406b250c1d86d329ce1fcd1c9"

# Defining a function to concatenate URL's and query strings easily

extend_URL <- function(URL, query_string){
    
    result <- paste(URL, "&", query_string, sep ="")
    return(result)
    
}

```

\newpage

# API Call for Genres

First, the list of official genre labels used in the website should be retrieved. 

```{r genre}

# URL template for the "genre" call

genre <- "https://api.themoviedb.org/3/genre/movie/list?api_key=<<api_key>>"

# Setting the language

genre <- extend_URL(genre, "language=en-US")

# Inserting the API key to the template

genre <- gsub("<<api_key>>", api_key, genre)

# Performing the call

genre_call <- httr::GET(genre)

# Checking the status of the call

genre_call$status_code

```

The status code implies that the call was successful. The next step is finding out the genre ID of "western".

\newpage

# API Call for Genres - Results

```{r id}

# Converting the call results to an easily readable list

genre_char <- base::rawToChar(genre_call$content)

genre_json <- jsonlite::fromJSON(genre_char, flatten = TRUE)

knitr::kable(genre_json$genres)

```

\newpage

# API Call for Movie Discovery

```{r discover}

# URL template for the "discover" call

discover <- "https://api.themoviedb.org/3/discover/movie?api_key=<<api_key>>"

# Setting the language

discover <- extend_URL(discover, "language=en-US")

# Retrieve the movies with highest number of votes

discover <- extend_URL(discover, "sort_by=vote_count.desc")

# Filtering by genre (Western's genre id is 37)

discover <- extend_URL(discover, "with_genres=37")

# Filtering by original language (Italian)

discover <- extend_URL(discover, "with_original_language=it")

# Inserting the API key to the template

discover <- gsub("<<api_key>>", api_key, discover)

# Performing the call

discover_call <- httr::GET(discover)

# Checking the status of the call

discover_call$status_code

```

The call was successful.

\newpage

# API Call for Movie Discovery - Results

```{r movies}

# Converting the call results to an easily readable list

discover_char <- base::rawToChar(discover_call$content)

discover_json <- jsonlite::fromJSON(discover_char, flatten = TRUE)

results <- discover_json$results

# Checking the retrieved field names

colnames(results)

```

\newpage

# Final Results

Finally, a clean summary table consisting of basic movie information can be generated as follows:

```{r final}

table <- results[c("title", "release_date", "vote_average", "vote_count")]

knitr::kable(table)

```

